---
title: "Overview of Extended Binary Operators for Data Manipulation"
author: "James J Balamuta"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of Extended Operators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

When writing functions or simply perform data analyses, sometimes you may 
wish for just that one additional operator or equality check that speeds up
the process greatly. The goal behind this package is to provide such
operators with as _minimal_ an overhead as possible.

# Package Installation

To install the package, please type:

```{r eval = FALSE}
install.packages("rops")
```

# Using the package

To use the package, please load it into the _R_ session by typing:

```{r}
library("rops")
```

# Coalescing Operators

## Null coalescing operator

```{r}
# Null value
x = NULL

# Before
y = if(is.null(x)){ "Unset" } else { x }

# After
y = x %??% "Unset"

y
```

## Missing value (`NA`) coalescing operator

To impute values when "missingness" is detected, the `ifna()` provides a
convenient interface.

```{r}
x = c(1:5, NA, NA, 8, NA, 10)
y = 1:10

# Before
ifelse(is.na(x), x, y)
ifelse(is.na(x), x, 3)

# After
ifna(x, y)
ifna(x, 3)
```

# Equality Operators

Checking for exact equivalence in object can be done using an infix operator
instead of relying upon `identical()`. 

**NB** There are cases where checking with `all.equal()` is preferred. 

```{r}
x = y = 1:5
x2 = x + 1

# Before
identical(x, y)
identical(x2, y)

# After
x %==% y
x2 %==% y

# Before
!identical(x, y)
!identical(x2, y)

# After
x %!=% y
x2 %!=% y
```


# Whole Numbers

Previously, using `is.integer()` would yield a check on the type of vector.
Here, the `is_whole()` function seeks to check the state of
all numbers individually in the vector regardless of whether they are
`numeric` or `integer`.

```{r}
x = c(1, 2, 3, 8.5)
y = c(1L, 2L, 3L)

# Before
is.integer(x)
is.integer(y)

# After
is_whole(x)
is_whole(y)
```

# Safe Sequence Generation

The colon operator (`:`) does not perform a check on whether the
increment should be positive or negative. This leads to issues when
iterating over an empty `vector` or `data.frame` using `1:length(obj)`. 
To avoid this, safe sequences `%:%` performs a check to make sure the
sequence is _positive_. If it is not, then an empty integer vector is
created to avoid having a loop run.

```{r}
x = NULL

# Before
1:length(x)

# After
1 %:% length(x)
```

# Not in Set

The not in set operator is a negation of the `%in%` operator that checks for
whether an element belongs to the set. 

```{r}
x = 1:5
set = 3:10

# Before
!(x %in% set)

# After
x %notin% set
```
